### 2022-11-14 definitions.tex
Update avec une preuve presque terminée

\textcolor{red}{Before defining WSTS, need a definition of TS and WQO}

@@

\textcolor{red}{Add lemma about infinite increasing subsequences ?}

@@

@@

\textcolor{red}{Expliquer ce que c'est une base d'abords.}

@@

\textcolor{red}{Un lemme je pense c'est important de le noter, peut-être pas comme ça peut être noter différement faudra voir}

@@

\begin{proof}
Assue we have a counter-example ...
\end{proof}

\textcolor{red}{Define WSTS}

\textcolor{red}{Define SWSTS - may be necessary}

@@

\textcolor{red}{For one-counter automata, in case the only tests are zero tests then 
I supposed $\leq$ is $\leq$ for non-zero integers, and I'll have to look-up/think for what to do with the zero element for instance. For TA it seem kind of nontrivial (since they allow $<c$ tests).}

\textcolor{red}{Define 'has effective pred-basis'. Maybe it should be included in WSTS definition, maybe it can be another def. I kind of like the idea of 'effective pred-basis' and
'decidable $\leq$' being independant from the WSTS definition}

@@

\textcolor{red}{Define what an Ideal is. Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion. Anti-ideal just downward closed so again just not that helpful a notation. Maybe have a }

\begin{definition}
A {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
$\uparrow I = I = \downarrow I$.
\end{definition}

A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
$s \in J$.

@@

\textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}

Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.

\textcolor{red}{Do we still demand this ?}

\begin{claim}{(stability of ideals)}
Let $I, J \subseteq S$ be upward-closed. Then the sets 
$\pred(I)$, $I \cup J$, and $I \cap J$ are upward-closed.
\end{claim}

@@

@@

\textcolor{red}{Seems like a reasonable assumption to me.}

@@

\textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}

>>>>

% \textcolor{red}{Before defining WSTS, need a definition of TS and WQO}

@@

% \textcolor{red}{Add lemma about infinite increasing subsequences ?}

\begin{lemma}
(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
\end{lemma}

\iffalse
\begin{proof}
Consider an infinite sequence and the set $M = \{i \in N \mid \forall j > i ~ x_i \not\leq x_j \}$. $M$ cannot
be infinite, otherwise it would lead to an infinite subsequence contradicting the wqo
hypothesis. Thus, $M$ is bounded and any $x_i$ with $i$ beyond $M$ can start an infinite
increasing subsequence.
\end{proof}
\fi

@@

A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \bigcup_{x \in I_b} \uparrow x$. 
% Higman investigated ordered sets with the finite basis property.

@@

% \textcolor{red}{Expliquer ce que c'est une base d'abords.}

@@

@@

\begin{proof}
Assue we have a counter-example.
We extract an infinite subsequence where
inclusion is strict: $I_{n_0} \subsetneq I_{n_1} \subsetneq I_{n_2} \cdots$. Now, for any $i>0$, we can pick some $x_i \in I_{n_i} \setminus I_{n_{i-1}}$. The well-quasi-ordering hypothesis means that the infinite sequence of $x_i$'s
contains an increasing pair $x_i \leq x_j$ for some $i<j$. Because $x_i$ belongs to an upward-
closed set $I_{n_i}$ we have $x_j \in I_{n_i} $, contradicting $x_j \not\in I_{n_{ j - 1}}$.
\end{proof}

% \textcolor{red}{Define WSTS}

% \textcolor{red}{Define SWSTS -  necessary}

@@

% \textcolor{red}{Define 'has effective pred-basis'. Maybe it should be included in WSTS definition, maybe it can be another def. I kind of like the idea of 'effective pred-basis' and 'decidable $\leq$' being independant from the WSTS definition}

\begin{proposition}
If $S$ is an WSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred^*(I )$ is upward-
closed.
\end{proposition}

\iffalse
Proof. Assume $s \in \pred^* (I )$. Then $s \rightarrow^* t$ for some $t \in I $. If now $s' \geq s$ then upward-compatibility entails that $s' \rightarrow^* t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred^*(I )$.
\fi

\begin{proposition}
If $S$ is an SWSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
\end{proposition}

\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi

@@

% \textcolor{red}{Define what an Ideal is. Actually an Ideal is just an upward-closed set, so maybe this just adds some confusion. Anti-ideal just downward closed so again just not that helpful a notation. Maybe have a }

\iffalse
\begin{definition}
A {\em bi-ideal} $I \subseteq S$ is an upward-closed and downward-closed set, i.e
$\uparrow I = I = \downarrow I$.
\end{definition}

@@

% \textcolor{red}{Probably one can already 'deduce' from this that ideal $I$ and anti-ideal $J$ for resp. good and bad states, in the case of e.g. timed automata would be given by sets of states}
\fi

A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
$s \in J$. 
Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.

% \textcolor{red}{Do we still demand this ?}

\begin{claim}{(stability of ideals)}
Let $I, J \subseteq S$ be upward-closed. Then the sets $I \cup J$, and $I \cap J$ are upward-closed.
\end{claim}

@@

\begin{fact}
% Fact 3 ([1]). 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,

(i) For every upward-closed set $I \subseteq S$, there exists a finite basis $B$ of $I$.

(ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
\end{fact}

\begin{definition}{ (index)}. 
For an upward-closed set $I \subseteq S$ and $k \geq 0$, let $I^k= \bigcup_{0 \leq j \leq k} pre^j(I)$.
The {\em index} $k(I)$ is the
smallest $k_0$ s.t. $I^k = I^{k_0}$ for all $k \geq k_0$.
\end{definition}

\begin{fact}
% Fact 4 (stop condition). 
Let $I \subseteq S$ be an upward-closed set and $k \geq 0$ s.t. $I^k = I^{k+1}$ , then $I^\ell = I^k$ for all $\ell \geq k$, i.e.,
$k(I) \leq k$. This also implies that $pre^* (I) = I^k $.
\end{fact}

\begin{lemma}
% Lemma 3 ([1]) 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
 Given a basis of an upward-closed set $I \subseteq S$, and a state $s$ of a strongly well-structured transition
system, we can decide whether we can reach $I$ from $s$.
\end{lemma}

\begin{proof}
We have to show that we can compute a basis of $I^{k+1}$ if we are given a basis of $I^k $. 
Then the
decidability of the stop condition follows directly. Let $B$ be a basis of $I^k$. 
We have
$$I^{k+1} = I \cup pre(I^k ) = I \cup
\bigcup_{s' \in B}
pre(\uparrow \{s' \}).$$

Since $pre(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $I^{k+1}$ . By
Fact 3, we can compute a basis of $I^{k+1}$.
\end{proof}

@@

% \textcolor{red}{Seems like a reasonable assumption to me.}

@@

% \textcolor{red}{In the Özkan paper the input include a basis of $\uparrow post^*(s)$. Not $100$\% sure it is necessary, think we can try to do without this assumption in the input.}

### 2022-11-14 proof.tex
Update avec une preuve presque terminée

\textcolor{red}{Pour l'instant je regarde la décidabilité du problème "symmétrique" c'est à dire que j'ai inversé toutes les propriétés et demandé à ce que SWSTS soit downward-compatible. Je pense ça doit être possible de faire les choses autrement ...}

@@

Assume  $(S, \rightarrow, \leq)$ is a SWSTS with downward compatibility, $I$ is a decidable downward-closed subset of $S$, and $J$ is an upward-closed set with a given basis.

We define inductively
$I^{k+1} = I \cup pre(I^k )$. Note that for all $k \in \N$, $I^k$ is downward-closed due to
the strongly downward compatibility of $(S, \rightarrow, \leq)$.

@@

In order to certify inclusion in $I^k$, we need to work with finite representations.
The next lemma uses that $I'$ and $J$ are downward- and upward-closed, respectively.

\begin{lemma}
Let $A \subseteq S$ be a set, $J \subseteq S$ upward-closed and $I' \subseteq S$ downward-closed. 
Then $A \cap J \subseteq I' \leftrightarrow (\downarrow  A) \cap J \subseteq I'$.
\end{lemma}

@@

\begin{corollary}
For all $k \in \N$,
$ post^*(s)\cap J \subseteq I^k \leftrightarrow (\downarrow  post^*(s)) \cap J \subseteq I^k$. 
% et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
\end{corollary}

@@

Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow Post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $post^*(s)$ if it contains $s$.
(on énumère des sur-approximations de la cloture par le bas de $post*(s)$ par leur bases finies).
Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\downarrow post^*(s)$ is such an inductive invariant and may
eventually be found.

Procedure 1 stops when it finds a basis $b(D)$ of an invariant $D$ such that
$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and we know the basis for $J$, we can
directly compute $b(D)  \cap J$.
% We can compute a basis of $I^{k+1}$ if we have a basis of $I^k$.
Due to the Lemma, 
$b(D)  \cap J \subseteq I^k$ implies
$D  \cap J \subseteq I^k$.
Hence
$b(D)  \cap J \subseteq I^k$ implies
$\downarrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
(since $D$ contains $post^*(s)$).

@@

\begin{figure}
\fbox{\parbox[t][3.cm][c]{6cm}{
$\phantom{a}$\\
(1)\phantom{aaaaa} $i \leftarrow 0$\\
(2)\phantom{aaaaa}\textbf{while} $\neg( \downarrow post^*(D_i) \subseteq D_i $
			 and $ s \in D_i$
			 and $ b(D)  \cap J \subseteq I^k  )$ \textbf{loop} \newline
	(3)\phantom{aaaaaa}$\phantom{aaaa} i \leftarrow i +1$ \newline
(4)\phantom{aaaaa}\textbf{end loop} \newline
(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
	}
}
	\caption{\textbf{Procedure 1:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
\end{figure}

@@

@@

>>>>

@@

\begin{proof}

Assume  $(S, \rightarrow, \leq)$ is a SWSTS with upward compatibility, $J$ is a decidable downward-closed subset of $S$, and $I$ is an upward-closed set with a given basis.

We define inductively
$I^{k} = \bigcup_{0 \leq j \leq k} pre^j(I)$. Note that for all $k \in \N$, $I^k$ is upward-closed due to
the strongly upward compatibility of $(S, \rightarrow, \leq)$.

@@

In order to certify inclusion in $I^k$, we need to work with finite representations.
The next lemma uses that $I$ and $J$ are upward- and downward-closed, respectively.

\begin{lemma}
Let $A \subseteq S$ be a set, $J \subseteq S$ downward-closed and $I \subseteq S$ upward-closed. 
Then $A \cap J \subseteq I \leftrightarrow (\uparrow  A) \cap J \subseteq I$.
\end{lemma}

@@

\begin{corollary}
For all $k \in \N$,
$ post^*(s)\cap J \subseteq I^k \leftrightarrow (\uparrow  post^*(s)) \cap J \subseteq I^k$. 
% et travailler avec $\downarrow post^*(s)$ à la place de $\uparrow post^*(s)$.
\end{corollary}

@@

Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\uparrow post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\uparrow post^*(s)$ if it contains $s$.
(on énumère des sur-approximations de la cloture par le haut de $post*(s)$ par leur bases finies).
Each “over-approximation” $D_i$ is given by its basis $b(D_i)$. Notice that, by standard monotonicity, $\uparrow post^*(s)$ is such an inductive invariant and may
eventually be found.

Procedure 1 stops when it finds a basis $b(D)$ of an invariant $D$ such that
$b(D)  \cap J \subseteq I^k$.  Since $b(D)$ is finite and $J$ is decidable, we can
directly compute $b(D)  \cap J$.
% We can compute a basis of $I^{k+1}$ if we have a basis of $I^k$.
We can compute a basis
of $I^{k+1}$ if we have a basis of $I^k$. %This follows by the proof of Lemma 3.
Due to the Lemma, 
$b(D)  \cap J \subseteq I^k$ implies
$D  \cap J \subseteq I^k$.
Hence
$b(D)  \cap J \subseteq I^k$ implies
$\uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$.
(since $D$ contains $ \uparrow post^*(s)$).

@@

\begin{figure}
\fbox{\parbox[t][3.cm][c]{6cm}{
$\phantom{a}$\\
(1)\phantom{aaaaa} $i \leftarrow 0$\\
(2)\phantom{aaaaa}\textbf{while} $\neg( \uparrow post^*(D_i) \subseteq D_i $
			 and $ s \in D_i$
			 and $ b(D)  \cap J \subseteq I^k  )$ \textbf{loop} \newline
	(3)\phantom{aaaaaa}$\phantom{aaaa} i \leftarrow i +1$ \newline
(4)\phantom{aaaaa}\textbf{end loop} \newline
(5)\phantom{aaaaa}\textbf{return} $\text{\textit{false}}$ \newline
	}
}
	\caption{\textbf{Procedure 1:} enumerates inductive invariants to find an inclusion certificate.}\label{procedure1}
\end{figure}

@@

\begin{enumerate}

\item $k$-resilience holds if, and only if, Procedure 1 terminates.

\item $k$-resilience do not hold if, and only if, Procedure 2 terminates.

\end{enumerate}

Proof:

\begin{enumerate}

\item By a simple induction, it can be shown that $\uparrow post^*(D) \subseteq D$ for every inductive invariant $D$. 
If Procedure 1 terminates, then
$post^*(s) \cap J \subseteq \uparrow post^*(s) \cap J \subseteq D  \cap J \subseteq I^k$
which implies that $k$-resilience holds.

It remains to show that Procedure 1 terminates whenever $k$-resilience holds. To do so, it suffices to prove that $\uparrow post^*(s)$ is an inductive invariant. Indeed, this implies that
$\uparrow post^*(s)$ is eventyally found by Procedure 1 when $k$-resilience holds.

Formally, let us show that $\uparrow post(\uparrow post^*(s)) \subseteq \uparrow post^*(s)$.
Let $b \in \uparrow post(\uparrow post^*(s))$ 
there exists $a', a, b$ such that
$s \rightarrow^* a'$,
$a' \leq a$,
$a \rightarrow b'$,
and
$b' \leq b$.
% By monotonicity, there exists b'' ≥ b' such that a → b'' . Therefore, x → ∗ b'' and b' ≥ b, hence b ∈ ↓ Post ∗ (x).
\textcolor{red}{By downward compatibility 
there exists $b'' \leq b'$ such that $a \rightarrow b'' $. Therefore, $x \rightarrow^* b''$ and $b' \geq b$, hence $b \in \downarrow post^*(x)$.}

\item Procedure 2 computes
 $post^{\leq n}(s) \cap J$
 until it finds an element not in $ I^k$.

If Procedure 2 terminates, then
$k$-resilience does not hold.
It remains to show that Procedure 2 terminates whenever $k$-resilience does not hold.

\end{enumerate}

@@

\begin{theorem}
{\sc SWSTS resilience} is decidable.
\end{theorem}

Iteratively check whether $k$-resilience holds. 
If this is the case, return $k_{min} = k$. Otherwise check whether 
$I^{k+1} = I^k$. If so, return $-1$ (false), otherwise
continue.
The stop condition is decidable
and by Fact 4 also sufficient. 
\end{proof}
